\section{Chord System Properties}
\label{sec:chord-sysprops}

The following properties are recognized by Chord.
The separator for list-valued properties can be either a blank space, a comma, a colon, or a semi-colon.

\begin{longtable}{|l|p{0.4in}|p{4.0in}|}
%\caption{TODO} \\
\hline
{\bf Property name} & {\bf Type} & {\bf Description and default value if any} \\
\endfirsthead
\hline
{\bf Property name} & {\bf Type} & {\bf Description and default value if any} \\
\endhead
\hline
\hline
\multicolumn{3}{|l|}{{\bf Chord resource properties}} \\
\verb+chord.main.dir+ & location & \verb+main+ directory in Chord's installation. \\
\hline
\verb+chord.lib.dir+ & location & Directory containing libraries needed by Chord.
\newline Default: \verb+${chord.main.dir}/lib/+ \\
\hline
\verb+chord.main.class.path+ & path & Classpath of Chord.
It includes the path specified by property \verb+chord.java.analysis.path+
to allow running user-defined program analyses.
\newline Default: See \verb+${chord.main.dir}/build.xml+ \\
\hline
\verb+chord.bddbddb.class.path+ & path & Classpath of \verb+bddbddb+.
\newline Default: See \verb+${chord.main.dir}/build.xml+ \\
\hline
\verb+chord.instr.agent.file+ & location &
Shared library implementing Chord instrumentation agent. \\
%\newline Default: \verb+${chord.main.dir}/lib/libchord_instr_agent.so+ \\
\hline
\verb+chord.javadoc.url+ & string & URL of the Javadoc location of program analyses.
It is used when publishing targets (i.e., when property \verb+chord.publish.targets+
is set to true).
\newline Default: \verb+http://chord.stanford.edu/javadoc/+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Chord boot properties}} \\
\hline
\verb+chord.work.dir+ & location & Working directory during Chord's execution.
\newline Default: current working directory \\
\hline
\verb+chord.props.file+ & location & Properties file loaded by \verb+${chord.main.dir}/build.xml+.
Any of the below properties, as well as other user-defined properties to be passed to Chord
(e.g., for user-defined analyses), may be set in this file.  Each relative (instead of absolute)
path element in the value of any property named \verb+chord.<...>.path+ set in this file is
converted to an absolute path element with respect to the directory containing this file.
\newline Default: \verb+${chord.work.dir}/chord.properties+ \\
\hline
\verb+chord.max.heap+ & string & Maximum memory size of JVM running Chord.
\newline Default: \verb+1024m+ \\
\hline 
\verb+chord.max.stack+ & string & Maximum thread stack size of JVM running Chord.
\newline Default: \verb+32m+ \\
\hline
\verb+chord.jvmargs+ & string & Arguments passed to JVM running Chord. \\
%\newline Default: \verb+"-showversion -ea -Xmx${chord.max.heap} -Xss${chord.max.stack}"+ \\
\hline
\verb+chord.bddbddb.max.heap+ & string & Maximum memory size of JVM running bddbddb.
\newline Default: \verb+1024m+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Chord task properties}} \\
\hline
\verb+chord.build.scope+ & bool & Compute analysis scope (i.e., reachable classes and methods)
using algorithm specified by properties \verb+chord.scope.kind+ and \verb+chord.reuse.scope+.
Program scope is computed regardless of the value of this property if another task (e.g., a
program analysis to be run) needs it.
\newline Default: \verb+false+ \\
\hline
\verb+chord.run.analyses+ & string list & List of names of program analyses to be run
in order.
\newline Default: \verb+""+ \\
\hline
\verb+chord.print.rels+ & string list & List of names of program relations whose contents
must be printed to files \verb+${chord.out.dir}/<...>.txt+ where \verb+<...>+ denotes the
relation name. 
\newline Default: \verb+""+ \\
\hline
\verb+chord.publish.targets+ & bool & Publish all targets defined by program analyses in
paths specified by properties \verb+chord.java.analysis.path+ and
\verb+chord.dlog.analysis.path+.
\newline Default: \verb+false+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Basic program properties}} \\
\hline
\verb+chord.main.class+ & class & Fully-qualified name of main class of given program
(e.g., \verb+com.example.Main+). \\
\hline
\verb+chord.class.path+ & path & Classpath of given program. \verb+${sun.boot.class.path}+
is implicitly appended to it.
\newline Default: \verb+${sun.boot.class.path}+ \\
\hline
\verb+chord.src.path+ & path & Java source path of given program.
\newline Default: \verb+""+ \\
\hline
\verb+chord.run.ids+ & string list & List of IDs to identify runs of given program.
\newline Default: \verb+0+ \\
\hline
\verb+chord.args.XXX+ & string & Command-line argument string to be used for given
program in run having ID \verb+XXX+.
\newline Default: \verb+""+ \\
\hline
\verb+chord.runtime.jvmargs+ & string & Arguments to JVM when running given program.
It is used when the program is executed to compute its scope (i.e., when property
\verb+chord.scope.kind+ is set to \verb+dynamic+) as well as when the instrumented
program is executed for dynamic analyses.
\newline Default: \verb+"-ea -Xmx1024m"+ \\ 
\hline
\hline
\multicolumn{3}{|l|}{{\bf Program scope properties}} \\ 
\hline
\verb+chord.scope.kind+ & [\verb+rta+ $|$ \verb+dynamic+] & Algorithm to compute
analysis scope (i.e., reachable classes and methods).  Current choices include
\verb+rta+ (static Rapid Type Analysis) and \verb+dynamic+ (dynamic analysis).
\newline Default: \verb+rta+ \\
\hline
\verb+chord.reuse.scope+ & bool & Treat analysis scope as the classes and methods listed
in files specified by properties \verb+chord.classes.file+ and \verb+chord.methods.file+,
respectively. Property \verb+chord.scope.kind+ is ignored if this property is set to
\verb+true+ and the two files exist.
\newline Default: \verb+false+ \\
\hline
\verb+chord.check.exclude+ & string list & List of prefixes of names of classes and
packages to be excluded by program analyses.
Interpretation of this property is analysis-specific. \\
%\newline Default: \verb+"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."+ \\
\hline
\verb+chord.instr.exclude+ & string list & List of prefixes of names of classes and
packages to be excluded from being instrumented.  Classes in package \verb+java.lang.ref+
are not instrumented regardless of the value of this property. \\
%\newline Default: \verb+"java.,sun.,com.sun.,com.ibm.,org.apache.harmony.,joeq.,jwutil."+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Program analysis properties}} \\
\hline
\verb+chord.java.analysis.path+ & path & Classpath containing program analyses written
in Java (i.e., \verb+@Chord+-annotated classes).
\newline Default: \verb+${chord.main.dir}/classes/+ \\
\hline
\verb+chord.dlog.analysis.path+ & path & Path of directories containing program analyses
written in Datalog (i.e., \verb+*.datalog+ and \verb+*.dlog+ files).
\newline Default: \verb+${chord.main.dir}/src/dlog/+ \\
\hline
\verb+chord.reuse.rels+ & bool & Construct program relations from BDDs stored on disk
(from a previous run of Chord) whenever possible instead of re-computing them.
\newline Default: \verb+false+ \\
\hline
\verb+chord.publish.results+ & bool & Publish the results of program analyses in HTML.
Interpretation of this property is analysis-specific.
\newline Default: \verb+true+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Program transformation properties}} \\
\hline
\verb+chord.ssa+ & bool & Do SSA transformation for all methods deemed reachable by the
algorithm used to compute analysis scope.
\newline Default: \verb+true+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Chord debug properties}} \\
\hline
\verb+chord.bddbddb.noisy+ & bool & Make bddbddb output debug info when solving Datalog programs.
\newline Default: \verb+false+ \\
\hline
\verb+chord.save.maps+ & bool & Write to file \verb+${chord.bddbddb.work.dir}/<...>.map+
when saving program domain named \verb+<...>+ (useful for debugging Datalog programs using
\verb+debug+ target of \verb+${chord.main.dir}/build.xml+).
\newline Default: \verb+true+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Chord instrumentation properties}} \\
\hline
\verb+chord.trace.pipe+ & bool & Implement the dynamic trace file as a POSIX pipe instead of a regular file.
\newline Default: \verb+true+ \\
\hline
\verb+chord.trace.block.size+ & int & Number of bytes to read/write in a single operation from/to the dynamic trace file.
\newline Default: \verb+4096+ \\
\hline
\verb+chord.runtime.class+ & class & Subclass of \verb+chord.project.Runtime+ used by dynamic program
analyses to handle events generated during an instrumented program's execution.
\newline Default: \verb+chord.project.BufferedRuntime+ \\
\hline
\hline
\multicolumn{3}{|l|}{{\bf Chord output properties}} \\
\hline
\verb+chord.out.dir+ & location & Directory to which Chord dumps all files.
\newline Default: \verb+${chord.work.dir}/chord_output/+ \\
\hline
\verb+chord.out.file+ & loc. & File to redirect standard output stream during Chord's execution.
\newline Default: \verb+${chord.out.dir}/log.txt+ \\
\hline
\verb+chord.err.file+ & loc. & File to redirect standard error stream during Chord's execution.
\newline Default: \verb+${chord.out.dir}/log.txt+ \\
\hline
\verb+chord.classes.file+ & loc. & File from/to which list of classes deemed reachable is read/written.
\newline Default: \verb+${chord.out.dir}/classes.txt+ \\
\hline
\verb+chord.methods.file+ & loc. & File from/to which list of methods deemed reachable is read/written.
\newline Default: \verb+${chord.out.dir}/methods.txt+ \\
\hline
\verb+chord.bddbddb.work.dir+ & loc. & Directory used by bddbddb as its input/output directory.
\newline Default: \verb+${chord.out.dir}/bddbddb/+ \\
\hline
\verb+chord.boot.classes.dir+ & loc. & Directory containing instrumented JDK classes used by given program.
\newline Default: \verb+${chord.out.dir}/boot_classes/+ \\
\hline
\verb+chord.user.classes.dir+ & loc. & Directory containing instrumented non-JDK classes of given program.
\newline Default: \verb+${chord.out.dir}/user_classes/+  \\
\hline
\verb+chord.instr.scheme.file+ & loc. & File specifying kind and format of events in dynamic trace file.
\newline Default: \verb+${chord.out.dir}/scheme.ser+ \\
\hline
\verb+chord.crude.trace.file+ & loc. & Crude dynamic trace file.
\newline Default: \verb+${chord.out.dir}/crude_trace.txt+ \\
\hline
\verb+chord.final.trace.file+ & loc. & Final dynamic trace file.
\newline Default: \verb+${chord.out.dir}/final_trace.txt+ \\
\hline
\end{longtable}
